---
title: "fig_1"
output: html_document
date: "2025-08-08"
editor_options: 
  chunk_output_type: console
---

```{r libraries}
library(pander)
library(knitr)

library(data.table)
setDTthreads(0)

library(stringr)

library(ggplot2)
library(viridis)
library(viridisLite)
```

```{r setup, include = FALSE}
setwd("/Users/hlange/Desktop/bioinfo/Flepseq_analysis")
knitr::opts_knit$set(root.dir = "/Users/hlange/Desktop/bioinfo/Flepseq_analysis",
                     base.dir = "/Users/hlange/Desktop/bioinfo/Flepseq_analysis")
```
```{r ticks}
#ticks and breaks for polyA profiles

custom_breaks <- c(0,10,20,30,40,50,60,70,80,90,100,110,120,130,140,150,160,170,180,190,200)
every_nth <- function(x, nth, empty = TRUE, inverse = FALSE) 
  {
  if (!inverse) {
    if(empty) {
      x[1:nth == 1] <- ""
      x
      } else {
        x[1:nth != 1]
        }
    } else {
      if(empty) {
        x[1:nth != 1] <- ""
        x
        } else {
          x[1:nth == 1]
        }
    }
}
```
```{r functions}
round_any = function(x, accuracy, f=ceiling){f(x/ accuracy) * accuracy}

# peak functions
# detect the first peak
find_first_peak <- function(x, y) {
    # Find local maxima
    peaks <- which(diff(sign(diff(y))) == -2) + 1
    
    # If no peaks are found, return NA
    if (length(peaks) == 0) {
        return(list(x = NA_real_, y = NA_real_))
    }
    
    # Select the first peak based on its position in the density (smallest x value)
    first_peak <- peaks[1]
    
    # Return the x and y value of the first peak
    return(list(x = x[first_peak], y = y[first_peak]))
}

# detect the most prominent peak in polyA size densities
find_single_peak <- function(x, y) {
  # Identify peaks
  peaks <- which(diff(sign(diff(y))) == -2) + 1
  if (length(peaks) == 0) {
    return(list(peak.x = NA_real_))
  }
  
  # Calculate prominences for each peak
  prominences <- y[peaks] - sapply(peaks, function(peak) {
    min(c(min(y[1:peak]), min(y[(peak + 1):length(y)])))
  })
  
  # Return x position of the most prominent peak as named list
  return(list(peak_x = x[peaks[which.max(prominences)]]))
}

# Function to calculate smoothed values
precalc_smooth <- function(data, x_col, y_col, group_cols, span = span) {
  smoothed_data <- C50[, {
    fit <- loess(get(y_col) ~ get(x_col), 
                 data= .SD, 
                 span = span, 
                 control = loess.control(surface = "direct"))  # Add this!
    smooth <- predict(fit)
    .(x = get(x_col), y = smooth)
  }, by = group_cols]
  return(smoothed_data)
}

# detect the most prom peak between 35 and 60
find_second_peak <- function(x, y) {
  # Restrict to the desired range
  in_range <- which(x >= 35 & x <= 60)
  if (length(in_range) < 3) {
    return(list(peak_x = NA_real_))
  }
  
  x_sub <- x[in_range]
  y_sub <- y[in_range]
  
  # Identify peaks: local maxima with a slope change
  peaks <- which(diff(sign(diff(y_sub))) == -2) + 1
  if (length(peaks) == 0) {
    return(list(peak_2 = NA_real_))
  }
  
  # Calculate prominences for each peak
  prominences <- sapply(peaks, function(peak) {
    left_min <- min(y_sub[1:peak])
    right_min <- min(y_sub[(peak + 1):length(y_sub)])
    y_sub[peak] - min(c(left_min, right_min))
  })
  
  # Find the most prominent peak
  most_prominent_index <- which.max(prominences)
  peak_2_value <- x_sub[peaks[most_prominent_index]]
  
  return(list(peak_2 =peak_2_value))
}
```
```{r font and fontsize}
my_font <- "Helvetica" 
axis_title_size<- 40 
axis_text_size <- 36 
legend_size <- 30
strip_text_size <- 36 
theme_set(
theme_classic(base_family = my_font, base_size = axis_text_size, base_line_size = 1) + 
  theme(
    axis.title = element_text(size = axis_title_size),
    panel.grid.major = element_line(linewidth = 0.3, color = "grey80"),
    panel.grid.minor.x = element_blank(),     # remove vertical minor grid lines
    panel.grid.minor.y = element_line(color = "grey85", linewidth = 0.3),  # keep horizontal minor grid
    axis.text = element_text(size = axis_text_size),
    legend.text = element_text(size = legend_size),
    legend.title = element_blank(),
    strip.text = element_text(size = strip_text_size, family = my_font),
    strip.background = element_blank(),
    axis.ticks = element_line(linewidth = 0.5)))



theme_no_grid <- function() {
  theme_classic(base_family = my_font, base_size = axis_text_size, base_line_size = 1) + 
  theme(
    axis.title = element_text(size = axis_title_size),
    #panel.grid.major = element_line(linewidth = 0.3, color = "grey80"),
    #panel.grid.minor.x = element_blank(),     # remove vertical minor grid lines
    #panel.grid.minor.y = element_line(color = "grey85", linewidth = 0.3),  # keep horizontal minor grid
    axis.text.x = element_blank(),
    legend.text = element_text(size = legend_size),
    legend.title = element_blank(),
    strip.text = element_text(size = strip_text_size, family = my_font),
    strip.background = element_blank(),
    axis.ticks = element_line(linewidth = 0.5))
}
```
```{r colors}

tissue_colors <-  c(  
   
   "seedlings" = "#004949",
   "rosette" = "#2F7F46")
  

# I use scale_color_viridis_d(option = "E", direction = -1, begin = 0.15, end = 0.8)+ for distributions split by values from other publications (i.e. te), and scale_color_viridis_d (option = "D", direction = -1, begin = 0.15, end = 0.8)+ for stacked barplots and distributions split by profile categories
```
```{r gag ten AGIs}
gag_ten <- c("AT1G79040",
             "AT1G29930",
             "AT1G61520",
             "AT1G20020",
             "AT2G39730",
             "AT2G42530",
             "AT2G38540",
             "AT2G21660",
             "AT5G01530",
             "AT5G02380")
```


# fread seedling
```{r fread}
seedling <- fread("/Users/hlange/Desktop/bioinfo/flepseq_preprocessing_paper/read_tables_paper/R10_seedling_WT_tailed.csv")
```
```{r C50 seedling}
grouped_counts <- seedling[, .N, by = .(AGI, rep)]
filtered_counts <- grouped_counts[N >= 50]
unique_combinations <- uniqueN(grouped_counts, by = "rep")
list_C50_seedling <- unique(filtered_counts[, if (.N == unique_combinations) .SD, by = AGI]$AGI)
print(length(list_C50_seedling))
rm(filtered_counts, unique_combinations)
C50 <- seedling[AGI %in% list_C50_seedling]
C50[, weight := 1/.N, by = .(AGI, rep)]
filtered_counts <- grouped_counts[N >= 150]
unique_combinations <- uniqueN(grouped_counts, by = "rep")
list_C150_seedling <- unique(filtered_counts[, if (.N == unique_combinations) .SD, by = AGI]$AGI)
rm(filtered_counts, grouped_counts, unique_combinations)
```
```{r table chack}
check_seedling <- seedling[, .(n=.N), by = rep]
AGI_seedling <- seedling[, .(Nr_AGI = uniqueN(AGI)), by = rep]
table <- merge(check_seedling, AGI_seedling, by = "rep")
```



# Table_AGI
## uridylation
```{r Uridylation levels }
Table_U <- C50[, .(n=.N), by = .(AGI, rep, tail, tissue)]

all_combinations <- CJ(
 tissue = unique(C50$tissue),
  AGI = unique(C50$AGI),
  rep = unique(C50$rep),
 tail = unique(C50$tail)
)
Table_U <- merge(all_combinations, Table_U, by = c("AGI", "rep","tail", "tissue"), all.x = TRUE) 
Table_U[, ':=' (tot = sum(n), perc_U = 100*n / sum(n)), by = .(AGI,rep, tissue)] 
Table_U <- Table_U[tail=="uridylated A-Tails"]
Table_U[is.na(perc_U), perc_U := 0]

Table_U[, mean_U := mean(perc_U, na.rm = TRUE), by = .(AGI, tissue)]
mean_Table_U <- unique(Table_U[, .(AGI, mean_U, tissue)], by = c("AGI", "tissue"))
mean_Table_U$mean_U <- round(mean_Table_U$mean_U, 2)
Table_U <- Table_U[, .(AGI, tissue, rep, perc_U)]
rm(all_combinations)
```
## percent short reads 
```{r short reads}
Table_short <- C50[, .(total_reads = .N, short_reads = sum(polya_size < 30, na.rm = TRUE), perc_short = 100 * sum(polya_size < 30, na.rm = TRUE) / .N
), by = .(AGI, rep, tissue)]
Table_short[, mean_short := mean(perc_short, na.rm = TRUE), by = .(AGI, tissue)]
mean_Table_short <- unique(Table_short[, .(AGI, mean_short, tissue)], by = c("AGI", "tissue"))
mean_Table_short$mean_short <- round(mean_Table_short$mean_short, 2)
Table_short <- Table_short[, .(AGI, tissue, rep, perc_short)]
```
## percent medium reads 
```{r mediumr reads}
Table_medium <- C50[, .(total_reads = .N, medium_reads = sum(polya_size >= 35 & polya_size <= 60 , na.rm = TRUE), perc_medium = 100 * sum(polya_size >= 35 & polya_size <= 60, na.rm = TRUE) / .N
), by = .(AGI, rep, tissue)]
Table_medium[, mean_medium := mean(perc_medium, na.rm = TRUE), by = .(AGI, tissue)]
mean_Table_medium <- unique(Table_medium[, .(AGI, mean_medium, tissue)], by = c("AGI", "tissue"))
mean_Table_medium$mean_medium <- round(mean_Table_medium$mean_medium, 2)
Table_medium <- Table_medium[, .(AGI, tissue, rep, perc_medium)]
```
## percent long reads
```{r long reads}
Table_long <- C50[, .(total_reads = .N, long_reads = sum(polya_size > 60 , na.rm = TRUE), perc_long = 100 * sum(polya_size > 60, na.rm = TRUE) / .N
), by = .(AGI, rep, tissue)]
Table_long[, mean_long := mean(perc_long, na.rm = TRUE), by = .(AGI, tissue)]
mean_Table_long <- unique(Table_long[, .(AGI, mean_long, tissue)], by = c("AGI", "tissue"))
mean_Table_long$mean_long <- round(mean_Table_long$mean_long, 2)
Table_long <- Table_long[, .(AGI, tissue, rep, perc_long)]
```
## RPM
```{r Table_reads}
Table_reads<- C50[, .(AGI_reads = .N),  by = .(AGI, rep, tissue)]
total <- C50[, .(total= .N), by = .(rep, tissue)]
Table_reads <- merge(Table_reads, total, by = c("rep", "tissue"))
Table_reads[, RPM := 1000000 *AGI_reads/total]
Table_reads[, mean_RPM := mean(RPM, na.rm = TRUE), by = .(AGI, tissue)]
mean_Table_reads <- unique(Table_reads[, .(AGI, mean_RPM, tissue)], by = c("AGI", "tissue"))
mean_Table_reads$mean_RPM <- round(mean_Table_reads$mean_RPM, 2)
Table_reads <- Table_reads[, .(AGI, tissue, rep, AGI_reads, RPM)]
```
## median poly a size
```{r Table_median}
Table_median <- C50[, .(median = median(polya_size)), by = .(AGI, rep, tissue)]
Table_median[, mean_median := mean(median, na.rm = TRUE), by = .(AGI, tissue)]
mean_Table_median <- unique(Table_median[, .(AGI, mean_median, tissue)], by = c("AGI", "tissue"))
mean_Table_median$mean_median <- round(mean_Table_median$mean_median, 2)
Table_median <- Table_median[, .(AGI, tissue, rep, median)]
```
## peaks per replicate
```{r density by Agi and rep}
DT_density <- C50[, .(density_data = list(density(polya_size, bw = "sj", adjust = 1))), by = .(AGI, tissue, rep)]

# Extract x and y from density_data list column
density_table <- DT_density[, {
  x <- density_data[[1]]$x
  y <- density_data[[1]]$y
  .(x, y)
}, by = .(AGI, tissue, rep)]

# calculate density and retrieve the results  to a table
all_1_pooled_peaks<- density_table[, {maxima <- find_first_peak(x, y)}, by = .(AGI, tissue, rep)]
all_1_pooled_peaks[, peak_1:= round(x, 0)]
all_1_pooled_peaks[,x := NULL]

all_p_pooled_peaks <- density_table[, {maxima <- find_single_peak(x, y)}, by = .(AGI, tissue, rep)]
all_p_pooled_peaks[, peak_p := round(peak_x, 0)]
all_p_pooled_peaks[, peak_x := NULL]

Table_peaks <- merge(all_1_pooled_peaks, all_p_pooled_peaks, by = c("AGI", "tissue","rep"))
Table_peaks <- Table_peaks[, .(AGI, tissue, rep, peak_1, peak_p)]
rm(all_1_pooled_peaks, all_p_pooled_peaks, DT_density, density_table)
```
## profile classification
```{r profile_1}
Table_peaks[, profile_1 := fcase(
  peak_p < 35, "short",
  peak_p >= 35 & peak_p <= 60, "inter",
  peak_p > 60, "long"
)]
Table_peaks[, profile_1 := factor(profile_1, levels  = c("short", "inter", "long"))]

```
## Table_AGI
```{r Table_AGI}

Table_AGI <- merge(Table_short, Table_medium, by = c("AGI", "tissue", "rep"))
Table_AGI <- merge(Table_AGI, Table_long, by = c("AGI", "tissue", "rep"))
Table_AGI <- merge(Table_AGI, Table_reads, by = c("AGI", "tissue", "rep"))
Table_AGI <- merge(Table_AGI, Table_median, by = c("AGI", "tissue", "rep"))
Table_AGI <- merge(Table_AGI, Table_peaks, by = c("AGI", "tissue", "rep"))
Table_AGI <- merge(Table_AGI, Table_U, by = c("AGI", "tissue", "rep"))
Table_AGI <-Table_AGI[, .(AGI, tissue, rep, median, perc_short, perc_medium, perc_long, perc_U, RPM, peak_1, peak_p,  profile_1)]
```
```{r write}
fwrite(Table_AGI, file = "R10_seedlings_WT_Table_AGI.csv")
```

# mean_table_AGI
```{r density by Agi}
DT_density <- C50[, .(density_data = list(density(polya_size, bw = "sj", adjust = 1))), by = .(AGI, tissue)]

# Extract x and y from density_data list column
density_table <- DT_density[, {
  x <- density_data[[1]]$x
  y <- density_data[[1]]$y
  .(x, y)
}, by = .(AGI, tissue)]

# calculate density and retrieve the results  to a table
all_1_pooled_peaks<- density_table[, {maxima <- find_first_peak(x, y)}, by = .(AGI, tissue)]
all_1_pooled_peaks[, peak_1:= round(x, 0)]
all_1_pooled_peaks[,x := NULL]

all_p_pooled_peaks <- density_table[, {maxima <- find_single_peak(x, y)}, by = .(AGI, tissue)]
all_p_pooled_peaks[, peak_p := round(peak_x, 0)]
all_p_pooled_peaks[, peak_x := NULL]

mean_Table_peaks <- merge(all_1_pooled_peaks, all_p_pooled_peaks, by = c("AGI", "tissue"))
mean_Table_peaks <- mean_Table_peaks[, .(AGI, tissue, peak_1, peak_p)]


rm(all_1_pooled_peaks, all_p_pooled_peaks, DT_density, density_table)
```
```{r profile_p}
mean_Table_peaks[, profile_p := fcase(
  peak_p < 35, "short",
  peak_p >= 35 & peak_p <= 60, "inter",
  peak_p > 60, "long"
)]
mean_Table_peaks[, profile_p := factor(profile_p, levels  = c("short", "inter", "long"))]
```
```{r mean Table_AGI}
mean_Table_AGI <- merge(mean_Table_short, mean_Table_medium, by = c("AGI", "tissue"))
mean_Table_AGI <- merge(mean_Table_AGI, mean_Table_long, by = c("AGI", "tissue"))
mean_Table_AGI <- merge(mean_Table_AGI, mean_Table_median, by = c("AGI", "tissue"))
mean_Table_AGI <- merge(mean_Table_AGI, mean_Table_reads, by = c("AGI", "tissue"))
mean_Table_AGI <- merge(mean_Table_AGI, mean_Table_peaks, by = c("AGI", "tissue"))
mean_Table_AGI <- merge(mean_Table_AGI, mean_Table_U, by = c("AGI", "tissue"))
mean_Table_AGI <- merge(mean_Table_AGI, mRNA_size, by = c("AGI"), all.y = FALSE)
```
```{r write}
fwrite(mean_Table_AGI, file = "R10_seedlings_WT_mean_Table_AGI.csv")
```
```{r top40}
ordered_AGIs <- mean_Table_AGI[order(-mean_RPM), AGI]
top40 <- ordered_AGIs[1:40]
ordered_AGIs <- mean_Table_AGI[order(mean_RPM), AGI]
low40 <- ordered_AGIs[1:40]
global_mean <- mean(mean_Table_AGI$mean_RPM)

closest40 <- mean_Table_AGI[
  order(abs(mean_RPM - global_mean)),
  AGI
][1:40]
```
```{r clean}
rm(mean_Table_long, mean_Table_short, mean_Table_median, mean_Table_medium, mean_Table_peaks, mean_Table_reads, mean_Table_U)
rm(Table_long, Table_median, Table_peaks, Table_reads, Table_short, Table_medium, Table_U)
rm(list = ls()[sapply(ls(), function(x) inherits(get(x), "gg"))])
rm(total, grouping_vars)
```


# poly a distri
```{r weighted distri}

all_combinations <- CJ(
  tissue = unique(C50$tissue),
  rep = unique(C50$rep),
  polya_size = unique(C50$polya_size))

weighted_data <-C50[, .(weighted_n = sum(weight)), by = .(tissue, polya_size, rep)]
weighted_data <- merge(all_combinations, weighted_data, by = c("tissue", "rep", "polya_size"), all.x = TRUE)
weighted_data <- weighted_data[is.na(weighted_n), weighted_n := 0]
weighted_data[, `:=` (total_weighted = sum(weighted_n), perc = 100 * (weighted_n) / sum(weighted_n)), by = .(tissue, rep)]
weighted_data[, mean:=mean(perc, na.rm = T), by = .(tissue, polya_size)]
weighted_data <- weighted_data[!is.na(mean)]
# Function to calculate smoothed values
precalc_smooth <- function(data, x_col, y_col, group_cols, span = 0.025) {
  smoothed_data <- data[, {
    fit <- loess(get(y_col) ~ get(x_col), 
                 data = .SD, 
                 span = span, 
                 control = loess.control(surface = "direct")) 
    smooth <- predict(fit)
    .(x = get(x_col), y = smooth)
  }, by = group_cols]
  return(smoothed_data)
}

grouping_vars <- c("tissue")  

weighted_data_smooth <- precalc_smooth(weighted_data, "polya_size", "mean", grouping_vars)
```
```{r pdf}
weighted <- ggplot() +
 geom_point(data= weighted_data, aes(x = polya_size, y= perc, color = tissue), size = 0.05)+
  geom_area(data = weighted_data_smooth, aes(x = x, y = y, fill = tissue), alpha = 0.3, position = "identity") +  

#  geom_vline(xintercept= c(20, 43), linetype = "dashed", color = "grey")+
  
  scale_fill_manual(values = tissue_colors)+
  scale_color_manual(values = tissue_colors)+
  scale_x_continuous(expand = c(0,0),
    limits = c(9, 200),
    breaks = custom_breaks,
    labels = every_nth(custom_breaks, 5, inverse = TRUE)) +
   scale_y_continuous(breaks = scales::pretty_breaks(n = 3))+     
  guides(fill = guide_legend(override.aes = list(size = 3, alpha = 0.4))) +

 
  theme(legend.position = "none")+
  
  labs(
  #title = "R10_stem_seedlings, changes of shoot short profiles",
   # subtitle = "lines at 20, 43",
    x = "polyA tail size",
    y = "frequency"
  )

weighted
ggsave(filename = "pdf_R10_seedlings_WT_C50_distri.pdf", weighted, width =8, height = 5, dpi = 1200)
```

```{r info profile }
info_profile <- Table_AGI[, .(AGI, rep, profile_1)]
C50p <- merge(C50, info_profile, by = c("AGI", "rep"))
```
```{r distri profile split}

all_combinations <- CJ(
  tissue = unique(C50p$tissue),
  rep = unique(C50p$rep),
  profile_1 = unique(C50p$profile_1),
  polya_size = unique(C50p$polya_size))

weighted_data <-C50p[, .(weighted_n = sum(weight)), by = .(tissue, polya_size, rep, profile_1)]
weighted_data <- merge(all_combinations, weighted_data, by = c("tissue", "rep", "polya_size", "profile_1"), all.x = TRUE)
weighted_data <- weighted_data[is.na(weighted_n), weighted_n := 0]
weighted_data[, `:=` (total_weighted = sum(weighted_n), perc = 100 * (weighted_n) / sum(weighted_n)), by = .(tissue, rep, profile_1)]
weighted_data[, mean:=mean(perc, na.rm = T), by = .(tissue, polya_size, profile_1)]
weighted_data <- weighted_data[!is.na(mean)]
# Function to calculate smoothed values
precalc_smooth <- function(data, x_col, y_col, group_cols, span = 0.025) {
  smoothed_data <- data[, {
    fit <- loess(get(y_col) ~ get(x_col), 
                 data = .SD, 
                 span = span, 
                 control = loess.control(surface = "direct")) 
    smooth <- predict(fit)
    .(x = get(x_col), y = smooth)
  }, by = group_cols]
  return(smoothed_data)
}

grouping_vars <- c("tissue", "profile_1")  

weighted_data_smooth <- precalc_smooth(weighted_data, "polya_size", "mean", grouping_vars)
```
```{r pdf}
weighted <- ggplot() +
 geom_point(data= weighted_data, aes(x = polya_size, y= perc, color = profile_1), size = 0.05)+
  geom_area(data = weighted_data_smooth, aes(x = x, y = y, fill = profile_1), alpha = 0.3, position = "identity") +  

#  geom_vline(xintercept= c(20, 43), linetype = "dashed", color = "grey")+
  
  scale_fill_viridis_d(option = "D", direction = -1, begin = 0.15, end = 0.8)+
  scale_color_viridis_d(option = "D", direction = -1, begin = 0.15, end = 0.8)+
  scale_x_continuous(expand = c(0,0),
    limits = c(9, 200),
    breaks = custom_breaks,
    labels = every_nth(custom_breaks, 5, inverse = TRUE)) +
   scale_y_continuous(breaks = scales::pretty_breaks(n = 3))+     
  guides(fill = guide_legend(override.aes = list(size = 3, alpha = 0.4))) +

 
  theme(legend.position = "none")+
  
  labs(
  #title = "R10_stem_seedlings, changes of shoot short profiles",
   # subtitle = "lines at 20, 43",
    x = "polyA tail size",
    y = "frequency"
  )

weighted
ggsave(filename = "pdf_R10_seedling_WT_C50_distri_profile_split.pdf", weighted, width =8, height = 5, dpi = 1200)
```

# poly a distri rosette
```{r fread}
rosette <- fread("/Users/hlange/Desktop/bioinfo/flepseq_preprocessing_paper/read_tables_paper/R10_rosette_WT_tailed.csv")
check <- rosette[, .(n=.N), by = .(tissue, rep, run)]
```
```{r C50 rosette}
grouped_counts <- rosette[, .N, by = .(AGI, rep)]
filtered_counts <- grouped_counts[N >= 50]
unique_combinations <- uniqueN(grouped_counts, by = "rep")
list_C50_rosette <- unique(filtered_counts[, if (.N == unique_combinations) .SD, by = AGI]$AGI)
print(length(list_C50_rosette))
rm(filtered_counts, unique_combinations)
C50r <- rosette[AGI %in% list_C50_rosette]
C50r[, weight := 1/.N, by = .(AGI, rep)]
filtered_counts <- grouped_counts[N >= 150]
unique_combinations <- uniqueN(grouped_counts, by = "rep")
list_C150_rosette <- unique(filtered_counts[, if (.N == unique_combinations) .SD, by = AGI]$AGI)
rm(filtered_counts, grouped_counts, unique_combinations)
```
```{r weighted distri}

all_combinations <- CJ(
  tissue = unique(C50r$tissue),
  rep = unique(C50r$rep),
  polya_size = unique(C50r$polya_size))

weighted_data <-C50r[, .(weighted_n = sum(weight)), by = .(tissue, polya_size, rep)]
weighted_data <- merge(all_combinations, weighted_data, by = c("tissue", "rep", "polya_size"), all.x = TRUE)
weighted_data <- weighted_data[is.na(weighted_n), weighted_n := 0]
weighted_data[, `:=` (total_weighted = sum(weighted_n), perc = 100 * (weighted_n) / sum(weighted_n)), by = .(tissue, rep)]
weighted_data[, mean:=mean(perc, na.rm = T), by = .(tissue, polya_size)]
weighted_data <- weighted_data[!is.na(mean)]
# Function to calculate smoothed values
precalc_smooth <- function(data, x_col, y_col, group_cols, span = 0.025) {
  smoothed_data <- data[, {
    fit <- loess(get(y_col) ~ get(x_col), 
                 data = .SD, 
                 span = span, 
                 control = loess.control(surface = "direct")) 
    smooth <- predict(fit)
    .(x = get(x_col), y = smooth)
  }, by = group_cols]
  return(smoothed_data)
}

grouping_vars <- c("tissue")  

weighted_data_smooth <- precalc_smooth(weighted_data, "polya_size", "mean", grouping_vars)
```
```{r pdf}
weighted <- ggplot() +
 geom_point(data= weighted_data, aes(x = polya_size, y= perc, color = tissue), size = 0.05)+
  geom_area(data = weighted_data_smooth, aes(x = x, y = y, fill = tissue), alpha = 0.3, position = "identity") +  

#  geom_vline(xintercept= c(20, 43), linetype = "dashed", color = "grey")+
  
  scale_fill_manual(values = tissue_colors)+
  scale_color_manual(values = tissue_colors)+
  scale_x_continuous(expand = c(0,0),
    limits = c(9, 200),
    breaks = custom_breaks,
    labels = every_nth(custom_breaks, 5, inverse = TRUE)) +
   scale_y_continuous(breaks = scales::pretty_breaks(n = 3))+     
  guides(fill = guide_legend(override.aes = list(size = 3, alpha = 0.4))) +

 
  theme(legend.position = "none")+
  
  labs(
  #title = "R10_stem_seedlings, changes of shoot short profiles",
   # subtitle = "lines at 20, 43",
    x = "polyA tail size",
    y = "frequency"
  )

weighted
ggsave(filename = "pdf_R10_rosette_WT_C50_distri.pdf", weighted, width =8, height = 5, dpi = 1200)
```

# barplot profile count seedlings
```{r barplot profile count C50}
Count <- Table_AGI[, .N, by = .(tissue, profile_1, rep)]
barplot<- ggplot(Count, aes(x = rep, y = N, fill = profile_1)) +
  geom_bar(stat = "identity", linewidth = 1.2) +
  facet_wrap(~tissue)+
  scale_fill_viridis_d(option = "D", direction = -1, begin = 0.15, end = 0.8)+
  #theme_classic(base_size = 18)+
  theme(legend.position = "right")+
  theme(axis.text.x = element_blank())+
 
  labs(

    y = "Nr AGIs",
    x = ""
  )

barplot
ggsave(filename ="pdf_R10_seedlings_barplot_profile_count.pdf", barplot, width = 6, height = 6, dpi = 1200)
```

# make quantiles from RPM
```{r match AGI}
RPM <- mean_Table_AGI[, .(AGI, mean_RPM)]
n_cuts <- 5  # You can change this to any integer ≥ 1

RPM[, RPM_quant := cut(mean_RPM, breaks = quantile(mean_RPM, probs = seq(0, 1, length.out = n_cuts + 1), na.rm = TRUE),  include.lowest = TRUE, labels = FALSE)]

quantile_limits <- quantile(
  RPM$mean_RPM,
  probs = seq(0, 1, length.out = n_cuts + 1),
  na.rm = TRUE
)
quantile_limits
```
quantile_limits
   0%       20%       40%       60%       80%      100% 
 12.780    23.890    36.354    59.982   125.436 35291.530 

```{r merge info to C50 and Table_AGI}

info_RPM <- RPM[, .(AGI, RPM_quant)]
info_RPM[, RPM_quant := as.character(RPM_quant)]
count <- info_RPM[, .(N=.N), by = "RPM_quant"]

C50_rpm <- merge(C50, info_RPM[,.(AGI, RPM_quant)], by = "AGI")
Table_AGI <- merge(Table_AGI, info_RPM, by = "AGI")
mean_Table_AGI <- merge(mean_Table_AGI, info_RPM, by = "AGI")
```
## weighted distri by RPM quantiles
```{r distri}
all_combinations <- CJ(
  rep = unique(C50_rpm$rep),
  RPM_quant = unique(C50_rpm$RPM_quant),
  polya_size = unique(C50_rpm$polya_size))

weighted_data <-C50_rpm[, .(weighted_n = sum(weight)), by = .(polya_size, rep,RPM_quant)]
weighted_data <- merge(all_combinations, weighted_data, by = c("rep", "polya_size", "RPM_quant"), all.x = TRUE)
weighted_data <- weighted_data[is.na(weighted_n), weighted_n := 0]
weighted_data[, `:=` (total_weighted = sum(weighted_n), perc = 100 * (weighted_n) / sum(weighted_n)), by = .(rep,RPM_quant)]
weighted_data[, mean:=mean(perc, na.rm = T), by = .(polya_size,RPM_quant)]

# Function to calculate smoothed values
precalc_smooth <- function(data, x_col, y_col, group_cols, span = 0.025) {
  smoothed_data <- data[, {
    fit <- loess(get(y_col) ~ get(x_col), 
                 data= .SD, 
                 span = span, 
                 control = loess.control(surface = "direct"))  # Add this!
    smooth <- predict(fit)
    .(x = get(x_col), y = smooth)
  }, by = group_cols]
  return(smoothed_data)
}

grouping_vars <- "RPM_quant" 

weighted_data_smooth <- precalc_smooth(weighted_data, "polya_size", "mean", grouping_vars)
```
```{r weighted distri RMP split}
weighted <- ggplot() +
 
  geom_area(data= weighted_data_smooth, aes(x = x, y = y, color = RPM_quant, fill =RPM_quant), alpha = 0.2, position = "identity") +  
  #geom_text(data= Count, aes(label = N), x = 160, y = 0.9, size = 6)+
  #geom_vline(xintercept= c(20, 43), linetype = "dashed", color = "red")+
  
 scale_fill_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
  scale_color_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
  scale_x_continuous(expand = c(0,0),
    limits = c(9, 200),
    breaks = custom_breaks,
    labels = every_nth(custom_breaks, 6, inverse = TRUE)
  ) +
  scale_y_continuous(limits = c(0, 1), breaks = scales::pretty_breaks(n = 3))+
  guides(fill = guide_legend(override.aes = list(size = 3, alpha = 0.4))) +
 
 
  theme(legend.position = "none")+
  
  labs(
    #title = "weighted distri",
    #subtitle = "Sorensen",
    x = "polyA tail size",
    y = "frequency"
  )

weighted
ggsave(filename = "pdf_R10_seedlings_WT_weighted_split_RPM_quant.pdf", weighted, width =8, height = 5, dpi = 1200)
```
```{r clean}
rm(weighted_data, weighted_data_smooth, all_combinations)
rm(list = ls()[sapply(ls(), function(x) inherits(get(x), "gg"))])
```


# make quantiles from RPM only upper 20 %
```{r match AGI}
RPM_2 <- mean_Table_AGI[mean_RPM > 125, .(AGI, mean_RPM)]
n_cuts <- 5  # You can change this to any integer ≥ 1

RPM_2[, RPM_quant_2 := cut(mean_RPM, breaks = quantile(mean_RPM, probs = seq(0, 1, length.out = n_cuts + 1), na.rm = TRUE),  include.lowest = TRUE, labels = FALSE)]

quantile_limits <- quantile(
  RPM_2$mean_RPM,
  probs = seq(0, 1, length.out = n_cuts + 1),
  na.rm = TRUE
)
quantile_limits
```
    0%       20%       40%       60%       80%      100% 
 125.100   158.744   212.154   300.094   551.922 35291.530 

```{r merge info to C50 and Table_AGI}

info_RPM <- RPM_2[, .(AGI, RPM_quant_2)]
info_RPM[, RPM_quant_2 := as.character(RPM_quant_2)]
count <- info_RPM[, .(N=.N), by = "RPM_quant_2"]

C50_2 <- merge(C50, info_RPM[,.(AGI, RPM_quant_2)], by = "AGI", all.y= FALSE)
Table_AGI_2 <- merge(Table_AGI, info_RPM, by = "AGI", all.y= FALSE)
mean_Table_AGI_2 <- merge(mean_Table_AGI, info_RPM, by = "AGI", all.y= FALSE)
```
## weighted distri by RPM quantiles
```{r distri}
all_combinations <- CJ(
  rep = unique(C50_2$rep),
  RPM_quant_2 = unique(C50_2$RPM_quant_2),
  polya_size = unique(C50_2$polya_size))

weighted_data <-C50_2[, .(weighted_n = sum(weight)), by = .(polya_size, rep,RPM_quant_2)]
weighted_data <- merge(all_combinations, weighted_data, by = c("rep", "polya_size", "RPM_quant_2"), all.x = TRUE)
weighted_data <- weighted_data[is.na(weighted_n), weighted_n := 0]
weighted_data[, `:=` (total_weighted = sum(weighted_n), perc = 100 * (weighted_n) / sum(weighted_n)), by = .(rep,RPM_quant_2)]
weighted_data[, mean:=mean(perc, na.rm = T), by = .(polya_size,RPM_quant_2)]

# Function to calculate smoothed values
precalc_smooth <- function(data, x_col, y_col, group_cols, span = 0.025) {
  smoothed_data <- data[, {
    fit <- loess(get(y_col) ~ get(x_col), 
                 data= .SD, 
                 span = span, 
                 control = loess.control(surface = "direct"))  # Add this!
    smooth <- predict(fit)
    .(x = get(x_col), y = smooth)
  }, by = group_cols]
  return(smoothed_data)
}

grouping_vars <- "RPM_quant_2" 

weighted_data_smooth <- precalc_smooth(weighted_data, "polya_size", "mean", grouping_vars)
```
```{r weighted distri RMP split}
weighted <- ggplot() +
 
  geom_area(data= weighted_data_smooth, aes(x = x, y = y, color = RPM_quant_2, fill =RPM_quant_2), alpha = 0.2, position = "identity") +  
  #geom_text(data= Count, aes(label = N), x = 160, y = 0.9, size = 6)+
  #geom_vline(xintercept= c(20, 43), linetype = "dashed", color = "red")+
  
 scale_fill_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
  scale_color_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
  scale_x_continuous(expand = c(0,0),
    limits = c(9, 200),
    breaks = custom_breaks,
    labels = every_nth(custom_breaks, 6, inverse = TRUE)
  ) +
  scale_y_continuous(limits = c(0, 1), breaks = scales::pretty_breaks(n = 3))+
  guides(fill = guide_legend(override.aes = list(size = 3, alpha = 0.4))) +
 
 
  theme(legend.position = "none")+
  
  labs(
    #title = "weighted distri",
    #subtitle = "Sorensen",
    x = "polyA tail size",
    y = "frequency"
  )

weighted
ggsave(filename = "pdf_R10_seedlings_WT_weighted_split_RPM_quant_2.pdf", weighted, width =8, height = 5, dpi = 1200)
```
```{r clean}
rm(weighted_data, weighted_data_smooth, all_combinations)
rm(list = ls()[sapply(ls(), function(x) inherits(get(x), "gg"))])
```

```{r box short}
box <- ggplot(Table_AGI_2, aes(x = rep, y = perc_short, color =RPM_quant_2)) +
  geom_boxplot(position = position_dodge(0.9), outlier.shape = NA,  fill = "white",  width=0.5, linewidth = 1, alpha=0.4, notch=F)+
  facet_wrap(~RPM_quant_2, nrow = 1)+
  scale_y_continuous(limits = c(0, 35))+
  scale_fill_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
  scale_color_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
   theme_no_grid()+
   theme(axis.title.x = element_blank())+
   theme (legend.position="none")+
  
labs(
     #x="",
     y= "% short reads")
box
ggsave(filename ="pdf_R10_seedlings_WT_box_short_RPM_split_2.pdf", box, width = 6, height = 6, dpi = 1200)

```
```{r box medium}
box <- ggplot(Table_AGI_2, aes(x = rep, y = perc_medium, color = RPM_quant_2)) +
  geom_boxplot(position = position_dodge(0.9), outlier.shape = NA,  fill = "white",  width=0.5, linewidth = 1, alpha=0.4, notch=F)+
  facet_wrap(~ RPM_quant_2, nrow = 1)+
  scale_y_continuous(limits = c(0, 35))+
  scale_fill_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
  scale_color_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
   theme_no_grid()+
   theme(axis.title.x = element_blank())+
   theme (legend.position="none")+
  
labs(
     #x="",
     y= "% medium")
box
ggsave(filename ="pdf_R10_seedlings_WT_box_medium_RPMsplit_2.pdf", box, width = 6, height = 6, dpi = 1200)

```



# half life Sorensen

```{r fread HLSorensen}

# formula to convert published alpha values to half-life in min: decay$half_life <- log(2)/decay$alpha_sov

HL_Sorensen_all <- fread("/Users/hlange/Desktop/bioinfo/published_data/Tables_pietro/Table_5Q_Sorenson.txt")
setnames(HL_Sorensen_all, "Half_life", "mean_HL")


plot <- ggplot(HL_Sorensen_all, aes(mean_HL))+
  geom_density()+
  geom_vline(xintercept = 1, linetype = "dashed", color = "red")+
  
 # scale_x_continuous(limits = c(0,3))+
  #scale_y_continuous(limits = c(0,3))+
  theme_classic(base_size = 15)+
  theme (legend.position="none")+

  labs(title = "HL_Sorensen_all.pdf")

  
plot
```
```{r match AGI}
HL_Sorensen <- unique(HL_Sorensen_all[AGI %in% Table_AGI$AGI], by = "AGI")
n_cuts <- 5  # You can change this to any integer ≥ 1

HL_Sorensen[, Sorensen := cut(mean_HL, breaks = quantile(mean_HL, probs = seq(0, 1, length.out = n_cuts + 1), na.rm = TRUE),  include.lowest = TRUE, labels = FALSE)]

quantile_limits <- quantile(
  HL_Sorensen$mean_HL,
  probs = seq(0, 1, length.out = n_cuts + 1),
  na.rm = TRUE
)
quantile_limits
HL_Sorensen[, Sorensen := as.character(Sorensen)]
HL_Sorensen <- merge(HL_Sorensen, mRNA_size, by = "AGI")
```

quantile_limits
         0%         20%         40%         60%         80%        100% 
   3.855295   57.672970   92.149672  144.921279  259.505169 1435.268596 
```{r get info}
info_Sorensen <- HL_Sorensen[, .(AGI, mean_HL, Sorensen)]
info_Sorensen[, Sorensen := as.character(Sorensen)]
count <- info_Sorensen[, .(N=.N), by = "Sorensen"]
```
```{r merge info_HL to C50 and Table_AGI}
C50_S <- merge(C50, info_Sorensen[,.(AGI, Sorensen)], by = "AGI", all.y = FALSE)
Table_AGI_S <- merge(Table_AGI, info_Sorensen, by = "AGI",  all.y = FALSE)
mean_Table_AGI_S <- merge(mean_Table_AGI, info_Sorensen, by = "AGI",  all.y = FALSE)
```


## weighted distri by Sorensen HL quantiles
```{r distri}
all_combinations <- CJ(
  rep = unique(C50_S$rep),
  Sorensen = unique(C50_S$Sorensen),
  polya_size = unique(C50_S$polya_size))

weighted_data <-C50_S[, .(weighted_n = sum(weight)), by = .(polya_size, rep,Sorensen)]
weighted_data <- merge(all_combinations, weighted_data, by = c("rep", "polya_size", "Sorensen"), all.x = TRUE)
weighted_data <- weighted_data[is.na(weighted_n), weighted_n := 0]
weighted_data[, `:=` (total_weighted = sum(weighted_n), perc = 100 * (weighted_n) / sum(weighted_n)), by = .(rep,Sorensen)]
weighted_data[, mean:=mean(perc, na.rm = T), by = .(polya_size,Sorensen)]

# Function to calculate smoothed values
precalc_smooth <- function(data, x_col, y_col, group_cols, span = 0.025) {
  smoothed_data <- data[, {
    fit <- loess(get(y_col) ~ get(x_col), 
                 data= .SD, 
                 span = span, 
                 control = loess.control(surface = "direct"))  # Add this!
    smooth <- predict(fit)
    .(x = get(x_col), y = smooth)
  }, by = group_cols]
  return(smoothed_data)
}

grouping_vars <- "Sorensen" 

weighted_data_smooth <- precalc_smooth(weighted_data, "polya_size", "mean", grouping_vars)
```

```{r weighted distri overlay}
weighted <- ggplot() +
 
  geom_area(data= weighted_data_smooth, aes(x = x, y = y, color =Sorensen, fill =Sorensen), alpha = 0.2, position = "identity") +  
  #geom_text(data= Count, aes(label = N), x = 160, y = 0.9, size = 6)+
  #geom_vline(xintercept= c(20, 43), linetype = "dashed", color = "red")+
  
 scale_fill_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
  scale_color_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
  scale_x_continuous(expand = c(0,0),
    limits = c(9, 200),
    breaks = custom_breaks,
    labels = every_nth(custom_breaks, 6, inverse = TRUE)
  ) +
  scale_y_continuous(limits = c(0, NA), breaks = scales::pretty_breaks(n = 3))+
  guides(fill = guide_legend(override.aes = list(size = 3, alpha = 0.4))) +
 
 
  theme(legend.position = "none")+
  
  labs(
    #title = "weighted distri",
    #subtitle = "Sorensen",
    x = "polyA tail size",
    y = "frequency"
  )

weighted
ggsave(filename = "pdf_R10_seedlings_WT_weighted_split_Sorensen.pdf", weighted, width =8, height = 5, dpi = 1200)
```
```{r clean}
rm(weighted_data, weighted_data_smooth, all_combinations)
rm(list = ls()[sapply(ls(), function(x) inherits(get(x), "gg"))])
```

```{r box short}
box <- ggplot(Table_AGI_S, aes(x = rep, y = perc_short, color =Sorensen)) +
  geom_boxplot(position = position_dodge(0.9), outlier.shape = NA,  fill = "white",  width=0.5, linewidth = 1, alpha=0.4, notch=F)+
  facet_wrap(~Sorensen, nrow = 1)+
  scale_y_continuous(limits = c(0, 35))+
  scale_fill_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
  scale_color_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
   theme_no_grid()+
   theme(axis.title.x = element_blank())+
   theme (legend.position="none")+
  
labs(
     #x="",
     y= "% short reads")
box
ggsave(filename ="pdf_R10_seedlings_WT_box_short_sorensen.pdf", box, width = 6, height = 6, dpi = 1200)

```
```{r box medium}
box <- ggplot(Table_AGI_S, aes(x = rep, y = perc_medium, color = Sorensen)) +
  geom_boxplot(position = position_dodge(0.9), outlier.shape = NA,  fill = "white",  width=0.5, linewidth = 1, alpha=0.4, notch=F)+
  facet_wrap(~ Sorensen, nrow = 1)+
  scale_y_continuous(limits = c(0, 35))+
  scale_fill_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
  scale_color_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
   theme_no_grid()+
   theme(axis.title.x = element_blank())+
   theme (legend.position="none")+
  
labs(
     #x="",
     y= "% medium")
box
ggsave(filename ="pdf_R10_seedlings_WT_box_medium_Sorensen.pdf", box, width = 6, height = 6, dpi = 1200)

```


## make quantiles from HL lower 20 %
```{r match AGI}
Sorensen_2 <- mean_Table_AGI_S[mean_HL < 57.672970, .(AGI, mean_HL)]
n_cuts <- 5  # You can change this to any integer ≥ 1

Sorensen_2[, Sorensen_2 := cut(mean_HL, breaks = quantile(mean_HL, probs = seq(0, 1, length.out = n_cuts + 1), na.rm = TRUE),  include.lowest = TRUE, labels = FALSE)]

quantile_limits <- quantile(
  Sorensen_2$mean_HL,
  probs = seq(0, 1, length.out = n_cuts + 1),
  na.rm = TRUE
)
quantile_limits
```
       0%       20%       40%       60%       80%      100% 
 3.855295 24.360361 33.694541 41.796053 50.177015 57.669696 

```{r merge info to C50 and Table_AGI}

info_Sorensen_2 <- Sorensen_2[, .(AGI, mean_HL, Sorensen_2)]
info_Sorensen_2[, Sorensen_2 := as.character(Sorensen_2)]
count <- info_Sorensen[, .(N=.N), by = "Sorensen_2"]

C50_S_2 <- merge(C50, info_Sorensen_2[,.(AGI, Sorensen_2)], by = "AGI", all.y= FALSE)
Table_AGI_S_2 <- merge(Table_AGI, info_Sorensen, by = "AGI", all.y= FALSE)
mean_Table_AGI_S_2 <- merge(mean_Table_AGI,info_Sorensen_2, by = "AGI", all.y= FALSE)
```
## weighted distri lower 20
```{r distri}
all_combinations <- CJ(
  rep = unique(C50_S_2$rep),
  Sorensen_2 = unique(C50_S_2$Sorensen_2),
  polya_size = unique(C50_S_2$polya_size))

weighted_data <-C50_S_2[, .(weighted_n = sum(weight)), by = .(polya_size, rep,Sorensen_2)]
weighted_data <- merge(all_combinations, weighted_data, by = c("rep", "polya_size", "Sorensen_2"), all.x = TRUE)
weighted_data <- weighted_data[is.na(weighted_n), weighted_n := 0]
weighted_data[, `:=` (total_weighted = sum(weighted_n), perc = 100 * (weighted_n) / sum(weighted_n)), by = .(rep,Sorensen_2)]
weighted_data[, mean:=mean(perc, na.rm = T), by = .(polya_size,Sorensen_2)]

# Function to calculate smoothed values
precalc_smooth <- function(data, x_col, y_col, group_cols, span = 0.03) {
  smoothed_data <- data[, {
    fit <- loess(get(y_col) ~ get(x_col), 
                 data= .SD, 
                 span = span, 
                 control = loess.control(surface = "direct"))  # Add this!
    smooth <- predict(fit)
    .(x = get(x_col), y = smooth)
  }, by = group_cols]
  return(smoothed_data)
}

grouping_vars <- "Sorensen_2" 

weighted_data_smooth <- precalc_smooth(weighted_data, "polya_size", "mean", grouping_vars)
```
```{r weighted distri 2nd HL split}
weighted <- ggplot() +
 
  geom_area(data= weighted_data_smooth, aes(x = x, y = y, color = Sorensen_2, fill =Sorensen_2), alpha = 0.2, position = "identity") +  
  #geom_text(data= Count, aes(label = N), x = 160, y = 0.9, size = 6)+
  #geom_vline(xintercept= c(20, 43), linetype = "dashed", color = "red")+
  
 scale_fill_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
  scale_color_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
  scale_x_continuous(expand = c(0,0),
    limits = c(9, 200),
    breaks = custom_breaks,
    labels = every_nth(custom_breaks, 6, inverse = TRUE)
  ) +
  scale_y_continuous(limits = c(0, 1.5), breaks = scales::pretty_breaks(n = 3))+
  guides(fill = guide_legend(override.aes = list(size = 3, alpha = 0.4))) +
 
 
  theme(legend.position = "none")+
  
  labs(
    #title = "weighted distri",
    #subtitle = "Sorensen",
    x = "polyA tail size",
    y = "frequency"
  )

weighted
ggsave(filename = "pdf_R10_seedlings_WT_weighted_split_Sorensen_2.pdf", weighted, width =8, height = 6, dpi = 1200)
```
```{r clean}
rm(weighted_data, weighted_data_smooth, all_combinations)
rm(list = ls()[sapply(ls(), function(x) inherits(get(x), "gg"))])
```


## box - und barplots
```{r box short}
box <- ggplot(Table_AGI_S_2, aes(x = rep, y = perc_short, color =Sorensen_2)) +
  geom_boxplot(position = position_dodge(0.9), outlier.shape = NA,  fill = "white",  width=0.5, linewidth = 1, alpha=0.4, notch=F)+
  facet_wrap(~Sorensen_2, nrow = 1)+
  scale_y_continuous(limits = c(0, 35))+
  scale_fill_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
  scale_color_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
   theme_no_grid()+
   theme(axis.title.x = element_blank())+
   theme (legend.position="none")+
  
labs(
     #x="",
     y= "% short reads")
box
ggsave(filename ="pdf_R10_seedlings_WT_box_short_Sorensen_2.pdf", box, width = 6, height = 6, dpi = 1200)

```
```{r box medium}
box <- ggplot(Table_AGI_S_2, aes(x = rep, y = perc_medium, color =Sorensen_2)) +
  geom_boxplot(position = position_dodge(0.9), outlier.shape = NA,  fill = "white",  width=0.5, linewidth = 1, alpha=0.4, notch=F)+
  facet_wrap(~Sorensen_2, nrow = 1)+
  scale_y_continuous(limits = c(0, 35))+
  scale_fill_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
  scale_color_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
   theme_no_grid()+
   theme(axis.title.x = element_blank())+
   theme (legend.position="none")+
  
labs(
     #x="",
     y= "% medium reads")
box
ggsave(filename ="pdf_R10_seedlings_WT_box_medium_Sorensen_2.pdf", box, width = 6, height = 6, dpi = 1200)

```


# TE Hsu

```{r fread TE Hsu}
TE_Hsu_all <- fread("/Users/hlange/Desktop/bioinfo/published_data/translation efficiency/TE_Hsu.csv")

plot <- ggplot(TE_Hsu_all, aes(mean_TE))+
  geom_density()+
  geom_vline(xintercept = 1, linetype = "dashed", color = "red")+
  
  scale_x_continuous(limits = c(0,3))+
  #scale_y_continuous(limits = c(0,3))+
  theme_classic(base_size = 15)+
  theme (legend.position="none")+

  labs(title = "TE_Hsu_all.pdf")

  
plot
#ggsave(filename ="pdf_R10_seedlings_WT_TE_all_Hsu.pdf", plot, width = 4, height = 2, dpi = 1200)
```
```{r match AGI}
TE_Hsu <- unique(TE_Hsu_all[AGI %in% Table_AGI$AGI], by = "AGI")
n_cuts <- 5  # You can change this to any integer ≥ 1

TE_Hsu[, Hsu := cut(mean_TE,breaks = quantile(mean_TE, probs = seq(0, 1, length.out = n_cuts + 1), na.rm = TRUE),  include.lowest = TRUE, labels = FALSE)]

quantile_limits <- quantile(
  TE_Hsu$mean_TE,
  probs = seq(0, 1, length.out = n_cuts + 1),
  na.rm = TRUE
)
quantile_limits
TE_Hsu [, Hsu := as.character(Hsu)]
TE_Hsu <- merge(TE_Hsu, mRNA_size, by = "AGI")
```
quantile_limits
         0%         20%         40%         60%         80%        100% 
 0.03693286  0.75122811  1.00662393  1.23730513  1.51920247 12.14984709 
```{r get info}
info_Hsu <- TE_Hsu[, .(AGI, mean_TE, Hsu)]
info_Hsu[, Hsu := as.character(Hsu)]
```
```{r merge info_TE to C50 and Table_AGI}
C50_H <- merge(C50, info_Hsu[,.(AGI, Hsu)], by = "AGI")
Table_AGI_H <- merge(Table_AGI, info_Hsu, by = "AGI")
mean_Table_AGI_H <- merge(mean_Table_AGI, info_Hsu, by = "AGI")
```

## weighted distri by TE quantiles
```{r distri}
all_combinations <- CJ(
  rep = unique(C50_H$rep),
  Hsu = unique(C50_H$Hsu),
  polya_size = unique(C50_H$polya_size))

weighted_data <-C50_H[, .(weighted_n = sum(weight)), by = .(polya_size, rep, Hsu)]
weighted_data <- merge(all_combinations, weighted_data, by = c("rep", "polya_size", "Hsu"), all.x = TRUE)
weighted_data <- weighted_data[is.na(weighted_n), weighted_n := 0]
weighted_data[, `:=` (total_weighted = sum(weighted_n), perc = 100 * (weighted_n) / sum(weighted_n)), by = .(rep, Hsu)]
weighted_data[, mean:=mean(perc, na.rm = T), by = .(polya_size, Hsu)]

# Function to calculate smoothed values
precalc_smooth <- function(data, x_col, y_col, group_cols, span = 0.025) {
  smoothed_data <- data[, {
    fit <- loess(get(y_col) ~ get(x_col), 
                 data= .SD, 
                 span = span, 
                 control = loess.control(surface = "direct"))  # Add this!
    smooth <- predict(fit)
    .(x = get(x_col), y = smooth)
  }, by = group_cols]
  return(smoothed_data)
}

grouping_vars <- "Hsu" 

weighted_data_smooth <- precalc_smooth(weighted_data, "polya_size", "mean", grouping_vars)
```
```{r weighted distri overlay}
weighted <- ggplot() +
 
  geom_area(data= weighted_data_smooth, aes(x = x, y = y, color = Hsu, fill = Hsu), alpha = 0.2, position = "identity") +  
  #geom_text(data= Count, aes(label = N), x = 160, y = 0.9, size = 6)+
  #geom_vline(xintercept= c(20, 43), linetype = "dashed", color = "red")+
  
 scale_fill_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
  scale_color_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
  scale_x_continuous(expand = c(0,0),
    limits = c(9, 200),
    breaks = custom_breaks,
    labels = every_nth(custom_breaks, 6, inverse = TRUE)
  ) +
  scale_y_continuous(limits = c(0, NA), breaks = scales::pretty_breaks(n = 3))+
  guides(fill = guide_legend(override.aes = list(size = 3, alpha = 0.4))) +
 
 
  theme(legend.position = "none")+
  
  labs(
    #title = "weighted distri",
    #subtitle = "Hsu",
    x = "polyA tail size",
    y = "frequency"
  )

weighted
ggsave(filename = "pdf_R10_seedlings_WT_weighted_split_Hsu.pdf", weighted, width =8, height = 5, dpi = 1200)
```
```{r clean}
rm(weighted_data, weighted_data_smooth, all_combinations)
rm(list = ls()[sapply(ls(), function(x) inherits(get(x), "gg"))])
```
## box - und barplots
```{r box short}
box <- ggplot(Table_AGI_H, aes(x = rep, y = perc_short, color =Hsu)) +
  geom_boxplot(position = position_dodge(0.9), outlier.shape = NA,  fill = "white",  width=0.5, linewidth = 1, alpha=0.4, notch=F)+
  facet_wrap(~Hsu, nrow = 1)+
  scale_y_continuous(limits = c(0, 35))+
  scale_fill_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
  scale_color_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
   theme_no_grid()+
   theme(axis.title.x = element_blank())+
   theme (legend.position="none")+
  
labs(
     #x="",
     y= "% short reads")
box
ggsave(filename ="pdf_R10_Heedlings_WT_box_Hhort_Hsu.pdf", box, width = 6, height = 6, dpi = 1200)

```
```{r box medium}
box <- ggplot(Table_AGI_H, aes(x = rep, y = perc_medium, color =Hsu)) +
  geom_boxplot(position = position_dodge(0.9), outlier.shape = NA,  fill = "white",  width=0.5, linewidth = 1, alpha=0.4, notch=F)+
  facet_wrap(~Hsu, nrow = 1)+
  scale_y_continuous(limits = c(0, 35))+
  scale_fill_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
  scale_color_viridis_d(option = "E", direction = 1, begin = 0.15, end = 0.8)+
   theme_no_grid()+
   theme(axis.title.x = element_blank())+
   theme (legend.position="none")+
  
labs(
     #x="",
     y= "% medium reads")
box
ggsave(filename ="pdf_R10_Heedlings_WT_box_medium_Hsu.pdf", box, width = 6, height = 6, dpi = 1200)

```


# dotplots black

```{r dotplot RPM vs median polyA size}
cor_test <- cor.test(mean_Table_AGI$mean_median, mean_Table_AGI$mean_RPM, method = "spearman")

corr_label <- paste0(
  "r = ", round(cor_test$estimate, 3),
  "\np = ", formatC(cor_test$p.value, format = "e", digits = 2)
)

dotplot<- ggplot()+
  geom_point(data = mean_Table_AGI, aes(x = mean_median, y = mean_RPM) , color = "black", size = 2, alpha = 0.2)+
  geom_smooth(data = mean_Table_AGI,aes(x = mean_median, y = mean_RPM),method = "lm", se = FALSE, color = "red", linewidth = 0.8 ) +
  annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.5,  label = corr_label,  size = 10) +
  scale_y_log10()+
  scale_x_continuous(limits = c(25, 200))+
  
  theme_no_grid()+
  theme(axis.text.x = element_text())+
 
  labs(
    x = "median polyA size",
    y = "RPM"
  )
dotplot
ggsave(filename ="pdf_R10_seedlings_WT_dotplot_RPM.pdf", dotplot, width = 8, height = 5, dpi = 1200)
```
```{r dotplot HL vs median polyA size}
cor_test <- cor.test(mean_Table_AGI_S$mean_median, mean_Table_AGI_S$mean_HL, method = "spearman")

corr_label <- paste0(
  "r = ", round(cor_test$estimate, 3),
  "\np = ", formatC(cor_test$p.value, format = "e", digits = 2)
)

dotplot<- ggplot()+
  geom_point(data = mean_Table_AGI_S, aes(x = mean_median, y = mean_HL) , color = "black", size = 2, alpha = 0.2)+
  geom_smooth(data = mean_Table_AGI_S,aes(x = mean_median, y = mean_HL),method = "lm", se = FALSE, color = "red", linewidth = 0.8 ) +
  annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.5,  label = corr_label,  size = 10) +
  
  scale_x_continuous(limits = c(25, 200))+
  theme_no_grid()+
  theme(axis.text.x = element_text())+
 
  labs(
    x = "median polyA size",
    y = "mRNA half-life"
  )
dotplot
ggsave(filename ="pdf_R10_seedlings_WT_dotplot_HL.pdf", dotplot, width = 8, height = 5, dpi = 1200)
```
```{r dotplot TE vs median polyA size}
cor_test <- cor.test(mean_Table_AGI_H$mean_median, mean_Table_AGI_H$mean_TE, method = "spearman")

corr_label <- paste0(
  "r = ", round(cor_test$estimate, 3),
  "\np = ", formatC(cor_test$p.value, format = "e", digits = 2)
)

dotplot<- ggplot()+
  geom_point(data = mean_Table_AGI_H, aes(x = mean_median, y = mean_TE) , color = "black", size = 2, alpha = 0.2)+
  geom_smooth(data = mean_Table_AGI_H,aes(x = mean_median, y = mean_TE),method = "lm", se = FALSE, color = "red", linewidth = 0.8 ) +
  annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.5,  label = corr_label,  size = 10) +
  scale_y_continuous(limits = c(0, 5))+
  scale_x_continuous(limits = c(25, 200))+
  theme_no_grid()+
  theme(axis.text.x = element_text())+
 
  labs(
    x = "median polyA size",
    y = "translation efficiency"
  )
dotplot
ggsave(filename ="pdf_R10_seedlings_WT_dotplot_mean_TE.pdf", dotplot, width = 8, height = 5, dpi = 1200)
```

# 40 most abundant in seedlings ordered by abundance
```{r merge info to mean table AGI}
mean_Table_AGI <- merge(mean_Table_AGI, info_Sorensen, by = "AGI",   all.x = TRUE, all.y = FALSE)
mean_Table_AGI <- merge(mean_Table_AGI, info_Hsu, by = "AGI",  all.x = TRUE, all.y = FALSE)
```
mean_Table_AGI
```{r}
ordered_AGIs <- mean_Table_AGI[order(-mean_RPM), AGI]
top40 <- ordered_AGIs[1:40]
reads_1 <- C50[AGI %in% top40]

reads_1[, AGI := factor(AGI, levels = ordered_AGIs)]
```
```{r}
table_1 <- Table_AGI[AGI %in% top40]
mean_table_1 <- mean_Table_AGI[AGI %in% top40]

```

```{r}
pRep_table <- function(dt, cols, round_col=NULL, digits=1) {
  out <- dt[AGI %in% top40, ..cols]
  if (!is.null(round_col)) out[, (round_col) := round(get(round_col), digits)]
  out
}
RPM <- pRep_table(mean_Table_AGI[AGI %in% top40], c("AGI","mean_RPM"), round_col="mean_RPM", digits=0)
median <- pRep_table(mean_Table_AGI[AGI %in% top40], c("AGI","mean_median"), round_col="mean_median", digits=0)
peak   <- pRep_table(mean_Table_AGI[AGI %in% top40], c("AGI","peak_p"), round_col="peak_p", digits=0)
HL     <- pRep_table(mean_Table_AGI[AGI %in% top40], c("AGI","mean_HL"), round_col="mean_HL", digits=0)
TE     <- pRep_table(mean_Table_AGI[AGI %in% top40], c("AGI","mean_TE"), round_col="mean_TE", digits=1)
```
```{r bin polyA}
reads_1[, polya_r := round_any(polya_size, 3)]

distri <- reads_1[, .(n=.N), by = .(AGI, polya_r, rep)]

all_combinations <- CJ(
  AGI       = top40,
  polya_r   = unique(reads_1$polya_r),
  rep       = unique(reads_1$rep)
)

distri <- merge(all_combinations, distri, by=c("AGI","polya_r","rep"), all.x=TRUE)
distri[is.na(n), n := 0]

distri[, tot := sum(n), by = .(AGI, rep)]
distri[, perc := 100*n/tot]

distri[, mean := mean(perc), by = .(AGI, polya_r)]

mean_distri <- unique(distri, by = c("AGI","polya_r"))

```
```{r}
# Apply order to all data tables used in the plots
for(dt in list(reads_1, distri, mean_distri, table_1, mean_table_1, peak, RPM, median, peak, HL, TE)) {
  dt[, AGI := factor(AGI, levels = ordered_AGIs)]
}

```

```{R distri pdf}
plot_di <- ggplot()+
  geom_col(data = mean_distri, aes(x=polya_r, y = mean), fill= "#004949", alpha = 0.5)+
  geom_point(data = distri, aes(x=polya_r, y = perc), color = "#004949", alpha = 0.8, size = 1)+
geom_text(data = peak,  aes(label = peak_p, x = Inf),  color = "magenta", y = Inf, vjust = 1.5, size = 12) +
geom_text(  data = RPM,  aes(label = mean_RPM, x = Inf),  hjust = 1, color = "black", y = Inf, vjust = 3, size = 12) +
geom_text(  data = HL,  aes(label = mean_HL, x = Inf),  hjust = 1, color = "#007A99", y = Inf, vjust = 4.5, size = 12) +
geom_text(  data = TE,  aes(label = mean_TE, x = Inf), hjust = 1, color = "#860086", y = Inf, vjust = 6, size = 12)+

  geom_vline(data = peak, aes(xintercept = peak_p), color = "magenta", linetype = "dashed", alpha= 1) +
  facet_wrap(~AGI, ncol =5)+
  
  #scale_fill_manual(values = tissue_colors, name = NULL)+
  #scale_color_manual(values =  tissue_colors, guide = "none")+
  scale_x_continuous(expand = c(0,0), limits=c(9, 200),
                     breaks = custom_breaks,
                     labels = every_nth(custom_breaks, 9, inverse = TRUE)) + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 3))+
  guides(fill = guide_legend(override.aes = list(size = 3, alpha = 1)))+
  ylab("density") +
  xlab("polyA tail size")+
 
  theme_no_grid()+
  theme (legend.position="none")+
 # theme(strip.text.x = element_blank())+
 # theme(strip.background = element_blank())+
  theme(axis.text.x = element_blank())+
  labs(#title = "R10_Genotype, always short",
       #subtitle = "binsize = 3",
  y= "frequency (%)",
  x= "polyA tail size")
plot_di

ggsave(filename ="pdf_R10_seedlings_WT_top40_ordered_distri.pdf", plot_di, width = 20, height = 40, dpi = 1200)

```


# 10 favourite in seedlings ordered by abundance
```{r}
gags_AGIs <- mean_Table_AGI[AGI %in% gag_ten]
ordered_AGIs <- gags_AGIs[order(-mean_RPM), AGI]
top10 <- ordered_AGIs[1:10]
reads_1 <- C50[AGI %in% top10]

reads_1[, AGI := factor(AGI, levels = ordered_AGIs)]
```
```{r}
table_1 <- Table_AGI[AGI %in% top10]
mean_table_1 <- mean_Table_AGI[AGI %in% top10]

```

```{r}
pRep_table <- function(dt, cols, round_col=NULL, digits=1) {
  out <- dt[AGI %in% top10, ..cols]
  if (!is.null(round_col)) out[, (round_col) := round(get(round_col), digits)]
  out
}

RPM_Rep     <- pRep_table(Table_AGI, c("AGI","RPM","rep"))
median_Rep  <- pRep_table(Table_AGI, c("AGI","median","rep"))
peak        <- pRep_table(mean_Table_AGI, c("AGI","peak_p"))

```
```{r bin polyA}
reads_1[, polya_r := round_any(polya_size, 3)]

distri <- reads_1[, .(n=.N), by = .(AGI, polya_r, rep)]

all_combinations <- CJ(
  AGI       = top10,
  polya_r   = unique(reads_1$polya_r),
  rep       = unique(reads_1$rep)
)

distri <- merge(all_combinations, distri, by=c("AGI","polya_r","rep"), all.x=TRUE)
distri[is.na(n), n := 0]

distri[, tot := sum(n), by = .(AGI, rep)]
distri[, perc := 100*n/tot]

distri[, mean := mean(perc), by = .(AGI, polya_r)]

mean_distri <- unique(distri, by = c("AGI","polya_r"))

```
```{r}
# Apply order to all data tables used in the plots
for(dt in list(reads_1, distri, mean_distri, table_1, mean_table_1, peak, RPM_Rep, median_Rep)) {
  dt[, AGI := factor(AGI, levels = ordered_AGIs)]
}

```

```{R distri pdf}
plot_di <- ggplot()+
  geom_col(data = mean_distri, aes(x=polya_r, y = mean), fill= "#004949", alpha = 0.5)+
 geom_point(data = distri, aes(x=polya_r, y = perc), color = "#004949", alpha = 0.8, size = 1)+
  geom_text(data = peak, aes(label = peak_p,  x = peak_p + 20), color = "#004949", y = Inf, vjust= 1.5, size = 12)+
 
 geom_vline(data = peak, aes(xintercept = peak_p), color = "magenta", linetype = "dashed", alpha= 1) +
  facet_wrap(~AGI, ncol =5)+
  
  #scale_fill_manual(values = tissue_colors, name = NULL)+
  #scale_color_manual(values =  tissue_colors, guide = "none")+
  scale_x_continuous(expand = c(0,0), limits=c(9, 200),
                     breaks = custom_breaks,
                     labels = every_nth(custom_breaks, 9, inverse = TRUE)) + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 3))+
  guides(fill = guide_legend(override.aes = list(size = 3, alpha = 1)))+
  ylab("density") +
  xlab("polyA tail size")+
 
  
  theme (legend.position="none")+
 # theme(strip.text.x = element_blank())+
 # theme(strip.background = element_blank())+
  theme(axis.text.x = element_blank())+
  labs(#title = "R10_Genotype, always short",
       #subtitle = "binsize = 3",
  y= "frequency (%)",
  x= "polyA tail size")
plot_di

ggsave(filename ="pdf_R10_seedlings_WT_top10_ordered_distri.pdf", plot_di, width = 20, height = 10, dpi = 1200)

```

# dotplots highlight
```{r dotplot RPM vs median polyA size}
dotplot<- ggplot()+
  geom_point(data = mean_Table_AGI, aes(x = mean_median, y = mean_RPM) , color = "black", size = 2, alpha = 0.2)+
  geom_point(data = mean_Table_AGI[AGI %in% top40], aes(x = mean_median, y = mean_RPM) , color = "red", size = 2, alpha = 0.8)+
  scale_y_log10()+
  scale_x_continuous(limits = c(25, 200))+
  theme_no_grid()+
  theme(axis.text.x = element_text())+
 
  labs(
    x = "median polyA size",
    y = "RPM"
  )
dotplot
ggsave(filename ="pdf_R10_seedlings_WT_dotplot_RPM_top40.pdf", dotplot, width = 8, height = 5, dpi = 1200)
```
```{r dotplot HL vs median polyA size}


dotplot<- ggplot()+
  geom_point(data = mean_Table_AGI_S, aes(x = mean_median, y = mean_HL) , color = "black", size = 2, alpha = 0.2)+
  geom_point(data = mean_Table_AGI_S[AGI %in% top40], aes(x = mean_median, y = mean_HL) , color = "red", size = 2, alpha = 0.8)+
  
  
  scale_x_continuous(limits = c(25, 200))+
  theme_no_grid()+
  theme(axis.text.x = element_text())+
 
  labs(
    x = "median polyA size",
    y = "mRNA half-life"
  )
dotplot
ggsave(filename ="pdf_R10_seedlings_WT_dotplot_HL_highlight.pdf", dotplot, width = 8, height = 5, dpi = 1200)
```
```{r dotplot TE vs median polyA size}

dotplot<- ggplot()+
  geom_point(data = mean_Table_AGI_H, aes(x = mean_median, y = mean_TE) , color = "black", size = 2, alpha = 0.2)+
  geom_point(data = mean_Table_AGI_H[AGI %in% top40], aes(x = mean_median, y = mean_TE) , color = "red", size = 2, alpha = 0.8)+
  scale_y_continuous(limits = c(0, 5))+
  scale_x_continuous(limits = c(25, 200))+
  theme_no_grid()+
  theme(axis.text.x = element_text())+
 
  labs(
    x = "median polyA size",
    y = "translation efficiency"
  )
dotplot
ggsave(filename ="pdf_R10_seedlings_WT_dotplot_mean_TE_highlight.pdf", dotplot, width = 8, height = 5, dpi = 1200)
```


